### 2020-10-01 latex/vlp.tex
Fixed errors that stopped compilation

\begin{thm}\label{thm:vlp-energy-stabilization}
   Take the VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
   \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
   Define \(\theta := \arctan_2(p,q)\). 
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \end{equation}
   with the property that the inequality is strict except at the
   coordinate axes.
   If the inequality is flipped, the VNHC dissipates energy. 
\end{thm}
\begin{proof}
   Let us examine the negative-time system
   \begin{equation}\label{eqn:negative-time-vlp}
      f(q,p) = 
      \begin{bmatrix}
         \dot{q} \\ \dot{p}
      \end{bmatrix} = \begin{bmatrix}
         -\frac{p}{ml(\theta)^2} \\
         mgl(\theta)\sin(q)
      \end{bmatrix}
      ,
   \end{equation}
   with \(D = (\mathcal{Q}\times \mathcal{P}) \backslash \{(\pm \pi, 0)\}\)
   the domain of interest on which we study this system.
   Choose, as a candidate Lyapunov-like function, the energy for the
   average-length pendulum with zero-potential at the bottom of the swing:
   \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
           + m g l_\text{avg} (1-\cos(q))
   \]
   This is positive definite at \((0,0)\) and has compact sublevel
   sets on \(D\). The derivative with respect to (\ref{eqn:negative-time-vlp})
   is
   \[
      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
              {l_\text{avg}^2l(\theta)^2}
      .
   \]
   The sign of \(\sin(q)p\) is dependent only on the quadrant in the
   \((q,p)\)-plane, so it is equal to the sign of \(\sin(2\theta)\). 
   Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
   \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
       .
   \]
   The derivative of \(E_\text{avg}\) is nonnegative since
   \begin{align*}
    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                &= \sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                &\leq 0 \text{ (by assumption)}
      ,
   \end{align*}
   which means \(E_\text{avg}\) is a Lyapunov function.
   Thus, the negative-time system is stable \cite{lyapunov}; all that remains is
   to show that (\ref{eqn:negative-time-vlp}) asymptotically converges to
   \((0,0)\). 
   For this, we will use a Krazovskii-LaSalle argument
   \cite{krazovskii_lasalle}.
   First, we define 
   \(Z = \left{(q,p) \in \mathbb{S}^1\times \R \mid 
      \dot{E}_\text{avg}(q,p) = 0 \right}\).
   It is easy to show that \(Z\) is the union of the coordinate axes
   \[
      Z= \mathbb{S}^1 \times \left{0\right} \cup 
         \left{0} \times \R\right}
      .
   \]
   Then, since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
   Krazovskii-LaSalle tells us that the origin is asymptotically stable on
   \(D\) for the negative-time system.
   Therefore, the orbit of the VNHC in the \((q,p)\) plane is
   diverging away from the origin in positive time, which means
   the VLP is gaining energy on average.
   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
   means \(E_\text{avg}\) is a Lyapunov function for the positive time system,
   so by the same argument the VLP is losing energy while converging
   asymptotically to \((0,0)\).
 \end{proof}

>>>>

\begin{thm}\label{thm:vlp-energy-stabilization}
   Take the VLP (\ref{eqn:vlp-hamiltonian}) with initial condition 
   \((q_0,p_0) \not \in \left\{(0,0),(\pm\pi,0)\right\}\).
   Define \(\theta := \arctan_2(p,q)\). 
   A VNHC of the form \(h(q,p) = l - l(\theta)\) injects energy if there exists 
   \(l_\text{avg} \in [\underline{l},\overline{l}]\) such that 
   \begin{equation}\label{eqn:vlp-energy-gain-condition}
      \left(l(\theta) - l_\text{avg}\right)sin(2\theta) \leq 0 \text{ }\forall \theta \in \mathbb{S}^1
   \end{equation}
   with the property that the inequality is strict except at the
   coordinate axes.
   If the inequality is flipped, the VNHC dissipates energy. 
\end{thm}
\begin{proof}
   Let us examine the negative-time system
   \begin{equation}\label{eqn:negative-time-vlp}
      f(q,p) = 
      \begin{bmatrix}
         \dot{q} \\ \dot{p}
      \end{bmatrix} = \begin{bmatrix}
         -\frac{p}{ml(\theta)^2} \\
         mgl(\theta)\sin(q)
      \end{bmatrix}
      ,
   \end{equation}
   with \(D = (\mathcal{Q}\times \mathcal{P}) \backslash \{(\pm \pi, 0)\}\)
   the domain of interest on which we study this system.
   Choose, as a candidate Lyapunov-like function, the energy for the
   average-length pendulum with zero-potential at the bottom of the swing:
   \[
       E_\text{avg}(q,p) := \frac{1}{2}\frac{p^2}{m l_\text{avg}^2} 
           + m g l_\text{avg} (1-\cos(q))
   \]
   This is positive definite at \((0,0)\) and has compact sublevel
   sets on \(D\). The derivative with respect to (\ref{eqn:negative-time-vlp})
   is
   \[
      \dot{E}_\text{avg} = \frac{g\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)}
              {l_\text{avg}^2l(\theta)^2}
      .
   \]
   The sign of \(\sin(q)p\) is dependent only on the quadrant in the
   \((q,p)\)-plane, so it is equal to the sign of \(\sin(2\theta)\). 
   Furthermore, by Lemma \ref{lemma:sign-of-cube} we have
   \[ 
       \sign{l(\theta)^3 - l_\text{avg}^3} = \sign{ l(\theta) - l_\text{avg}}
       .
   \]
   The derivative of \(E_\text{avg}\) is nonnegative since
   \begin{align*}
    \sign{\dot{E}_\text{avg}} &= \sign{\sin(q)p \left(l(\theta)^3 - l_\text{avg}^3\right)} \\
                &= \sign{\sin(2\theta) \left(l(\theta) - l_\text{avg}\right)} \\
                &\leq 0 \text{ (by assumption)}
      ,
   \end{align*}
   which means \(E_\text{avg}\) is a Lyapunov function.
   Thus, the negative-time system is stable \cite{lyapunov}; all that remains is
   to show that (\ref{eqn:negative-time-vlp}) asymptotically converges to
   \((0,0)\). 
   For this, we will use a Krazovskii-LaSalle argument
   \cite{krazovskii_lasalle}.
   First, we define 
   \[
      Z = \left\{(q,p) \in \mathbb{S}^1\times \R \mid 
      \dot{E}_\text{avg}(q,p) = 0 \right\}
      .
   \]
   It is easy to show that \(Z\) is the union of the coordinate axes
   \[
      Z= \mathbb{S}^1 \times \left\{0\right\} \cup 
         \left\{0\right\} \times \R
      .
   \]
   Krazovskii-LaSalle tells us solutions to the negative-time ODE converge to
   the largest invariant set in \(Z\).
   Since \(f(q,p) = (0,0)\) only when \((q,p) = (0,0)\),
   the origin is asymptotically stable on \(D\) for this negative-time system.
   Therefore, the positive-time system is diverging away from the origin 
   in the \((q,p)\) plane, which means the VLP is gaining energy on average.
   Flipping the inequality of (\ref{eqn:vlp-energy-gain-condition}) 
   means \(E_\text{avg}\) is a Lyapunov function for the positive time system,
   so by the same argument the VLP is losing energy while converging
   asymptotically to \((0,0)\).
 \end{proof}
